Counterexample to $1/T$ convergence of herding in infinite dimensions?

### 2013-12-04 nips-poster/nips-poster.tex
add sections I was responsible for; minor edits elsewhere

\begin{frame}{} 
%\titlepage  
%\vfill
\begin{columns}
\begin{column}{0.48\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 1		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Motivation}
Suppose we want to solve the following \emph{saddle point} problem:
\[ \min_u \max_{\theta} L(u, \theta), \]
where $L(u,\theta) = h(u) + \langle u, \theta \rangle - R(\theta)$. 
We assume that $h$ and $R$ are both convex and that $\langle \cdot, \cdot \rangle$ 
is an arbitrary bilinear map.

\textbf{Intuition:} $u$ is the ``primal'', $\theta$ is the ``dual''.

Tie-in with optimization: can recast as minimizing
\[ L(u) \eqdef \max_{\theta} L(u, \theta) = h(u) + R^*(u). \]

@@

\begin{block}{\large Fenchel conjugate properties}
Fenchel conjugates are a key component of \emph{Legendre-Fenchel duality} 
and also show up in mirror descent. It turns out that they will also be 
important for generalizing the Frank-Wolfe algorithm.

\textbf{Property 1:} If $R$ is convex then $R^{**} = R$.

\textbf{Property 2:} $\partial R^*(u) = \argmin_{\theta} u^T\theta - R(\theta)$.
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Being (Too) Greedy}
JHH
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Being Greedy Enough: Boosted Mirror Descent}
JHH

This block would contain the statement of the algorithms and the geometric intuition 
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{column}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 2		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{column}{0.48\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN APPLICATIONS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Special Cases and Applications}
Many particular algorithms can be cast into our framework.
\begin{columns}[t]
\begin{column}{0.45\linewidth}

@@

\end{column}
\begin{column}{0.45\linewidth}
\begin{block}{AHK low-rank SDP solver}
Suppose we want to solve the SDP
\begin{align*}
\text{maximize }   & \Tr(A^TX) \\
\text{subject to } & X \succeq 0 \\
                  & X_{ii} \leq 1 \, \forall i
\end{align*}
Set \small{$L(X,y) = \Tr(A^TX) + \sum_{i=1}^n \left[y_i(X_{ii}-1) - \eta y_i\log y_i\right]$}.
Then updating $X$ is an eigenvalue problem and each update is rank $1$. This 
yields a variant of the Arora-Hazan-Kale fast SDP solver.
\end{block}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Convergence of Boosted Mirror Descent}
\begin{columns}[t]
\begin{column}{0.45\linewidth}
\begin{theorem}
Consider the primal (respectively the dual) algorithm. Suppose that $h$ ($R$) is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$ ($r = \sup_{u} \|u\|_{*}$). Then
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\end{column}
\begin{column}{0.45\linewidth}
\begin{corollary} 
Under the hypotheses of the Theorem, for $\alpha_{t} = 1$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
and for $\alpha_t = t$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
\end{corollary}
\end{column}
\end{columns}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\small Acknowledgements}
{\footnotesize Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
to the paper. JS was supported by the Hertz Foundation. JHH was supported by the U.S. Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\small References}
\begin{footnotesize}
\bibliographystyle{plainnat}
\bibliography{herding}
\end{footnotesize}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{frame}{} 
%\titlepage  
%\vfill
\begin{columns}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 1		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{column}{0.46\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Motivation}
Suppose we want to solve the following \emph{saddle point} problem:
\[ \min_u \max_{\theta} L(u, \theta), \]
where $L(u,\theta) = h(u) + \langle u, \theta \rangle - R(\theta)$. 
We assume that $h$ and $R$ are both convex and that $\langle \cdot, \cdot \rangle$ 
is an arbitrary bilinear map.
\newline\newline
\textbf{Intuition:} $u$ is the ``primal'', $\theta$ is the ``dual''.
\newline\newline
\textbf{Tie-in with optimization:} can recast as minimizing
\[ L(u) \eqdef \max_{\theta} L(u, \theta) = h(u) + R^*(u). \]

@@

\begin{block}{\large Fenchel conjugate properties}
Fenchel conjugates are a key component of \emph{Legendre-Fenchel duality} 
and also show up in mirror descent. It turns out that they will also be 
important for generalizing the Frank-Wolfe algorithm.
\newline\newline
\textbf{Property 1:} If $R$ is convex then $R^{**} = R$.
\qquad\qquad
\textbf{Property 2:} $\partial R^*(u) = \argmin_{\theta} u^T\theta - R(\theta)$.
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Being (Too) Greedy}
\begin{columns}[t]
\begin{column}{.48\linewidth}
We might try the following ``iterative best response'' updates:
\begin{align*}
u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\theta_t) \\
\theta_{t+1}   &= \argmax_{\theta} L(u_t, \theta)
\end{align*}
\end{column}
\begin{column}{.48\linewidth}
{\bf Problem:} let $u,\theta \in \bR$, $h(u) = \frac{1}{2}u^2$, $R(\theta) = \frac{1}{2}\theta^2$. 
Then:
\begin{align*}
u_t\,\,\,\,\,\, &= \argmin_u \left[\frac{1}{2}u^2 + u\theta\right]             = -\theta_t \\
\theta_{t+1}    &= \argmax_{\theta} \left[u\theta - \frac{1}{2}\theta^2\right] = u_t
\end{align*}
\begin{center}
\textcolor{red}{OSCILLATION}
\end{center}
\end{column}
\end{columns}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{columns}[t]
\begin{column}{.43\linewidth}
\begin{block}{\large Being (Less) Greedy: \\ Boosted Mirror Descent}
Replace $u_t$ with a weighted average
\begin{align*}
\hat{u}_t &\eqdef \frac{1}{A_{t}} \sum_{s=1}^t \alpha_{s} u_s & A_{t} = \sum_{s=1}^{t} \alpha_{s}
\end{align*}
New updates (``dual'' algorithm):
\begin{align*}
u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\theta_t)  = && \partial h^{*}(-\theta_{t}) \\
\theta_{t+1}    &= \argmax_{\theta} L(\hat{u}_t, \theta)  = &&\partial R^{*}(\hat u_{t})
\end{align*}
Alternatively, replace $\theta_{t}$ with a weighted average (``primal'' algorithm):
\begin{align*}
u_t\,\,\,\,\,\, &= \argmin_{u} L(u,\hat\theta_t) = && \partial h^{*}(-\hat\theta_{t}) \\
\theta_{t+1}    &= \argmax_{\theta} L(u_t, \theta) = &&\partial R^{*}(u_{t})
\end{align*}
\end{block}
\end{column}
\begin{column}{.53\linewidth}
\begin{block}{\large A Geometric Picture (Primal)}
\begin{center}
\includegraphics[width=10in]{figures/primal-demo.pdf}
\end{center}
For the primal algorithm,
\begin{align*}
u_t\,\,\,\,\,\, &= \partial h^{*}(-\hat\theta_{t}) \text{ is a point where $-\hat\theta_{t} \in \partial h$} \\ 
\theta_{t+1}    &= \partial R^{*}(u_{t}) \text{ is the gradient of $R^{*}$ at $u_{t}$}
\end{align*}
So $u_{t}$ is chosen such that the gradient contributed by $h$ at $u_{t}$ cancels out the gradient contributed by $R^{*}$. 
\end{block}
\end{column}
\end{columns}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{column}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 2		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{column}{0.50\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN APPLICATIONS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Special Cases and Applications}
Many particular algorithms can be cast into our framework.
\begin{columns}[t]
\begin{column}{0.01\linewidth}\end{column}
\begin{column}{0.47\linewidth}

@@

\end{column}
\begin{column}{0.47\linewidth}
\begin{block}{AHK low-rank SDP solver}
Suppose we want to solve the SDP
\begin{align*}
\text{maximize }   & \Tr(A^TX) \\
\text{subject to } & X \succeq 0 \\
                  & X_{ii} \leq 1 \, \forall i
\end{align*}
Set \small{$L(X,y) = \Tr(A^TX) + \sum_{i=1}^n \left[y_i(X_{ii}-1) - \eta y_i\log y_i\right]$}.
Then updating $X$ is an eigenvalue problem and each update is rank $1$. This 
yields a variant of the Arora-Hazan-Kale fast SDP solver.
\end{block}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Convergence of Boosted Mirror Descent}
\begin{columns}[t]
\begin{column}{0.01\linewidth}\end{column}
\begin{column}{0.47\linewidth}
\begin{theorem}
Consider the primal (respectively the dual) algorithm. Suppose that $h$ ($R$) is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$ ($r = \sup_{u} \|u\|_{*}$). Then
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\end{column}
\begin{column}{0.47\linewidth}
\begin{corollary} 
Under the hypotheses of the Theorem, for $\alpha_{t} = 1$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
and for $\alpha_t = t$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
\end{corollary}
\end{column}
\end{columns}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\small Acknowledgements}
{\tiny Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
for helpful discussions. JS was supported by the Hertz Foundation. JHH was supported by the U.S. Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BEGIN BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{block}{\small References}
%\begin{footnotesize}
%\bibliographystyle{plainnat}
%\bibliography{herding}
%\end{footnotesize}
%\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% END BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
